LEGWORK: A python package for computing the evolution and detectability \\ of stellar-origin gravitational-wave sources with space-based detectors

### 2021-11-15 src/ms.tex
remove anchors from links, fix typos

% LEGWORK stuff
\newcommand{\lw}{LEGWORK}
\newcommand{\lwColour}{SeaGreen}
\newcommand{\lwModURL}[1]{https://legwork.readthedocs.io/en/latest/modules.html\#module-legwork.#1}
\newcommand{\lwModLink}[1]{\href{https://legwork.readthedocs.io/en/latest/modules.html\#module-legwork.#1}{\color{\lwColour}{\texttt{#1}}}}
\newcommand{\lwFuncURL}[2]{https://legwork.readthedocs.io/en/latest/api/legwork.#1.#2.html\#legwork.#1.#2}

@@

It is often convenient to convert between orbital frequency, $f_{\rm orb}$, and the semi-major axis, $a$, of a binary and this can be accomplished with Kepler's third law.
\begin{equation}{\lwFuncURL{utils}{get_a_from_f_orb}}
    a = \qty(\frac{G(m_1 + m_2)}{(2 \pi f_{\rm orb})^{2}})^{1/3},
    \label{eq:kepler3rd}
\end{equation}
where $G$ is the gravitational constant. Inversely we could write
\begin{equation}{\lwFuncURL{utils}{get_a_from_f_orb}}
    f_{\rm orb} = \frac{1}{2 \pi} \sqrt{\frac{G(m_1 + m_2)}{a^3}}.
    \label{eq:kepler3rd_reverse}
\end{equation}

For circular binaries, gravitational-wave emission occurs at twice the orbital frequency ($f_{\rm GW} = 2 f_{\rm orb}$). However, for eccentric binaries, we need to consider all frequency harmonics of gravitational-wave emission. These are defined such that the $n^{\rm th}$ harmonic frequency is
\begin{equation}{}
    f_{n} \equiv n \cdot f_{\rm orb}.
    \label{eq:nth_freq}
\end{equation}
It will be important to know the relative gravitational-wave power radiated into the $n^{\rm th}$ harmonic for a binary with eccentricity $e$ for the following strain and SNR calculations. This is given by \citet[][Eq.\,20]{Peters+1963}
\begin{equation}{\lwFuncURL{utils}{peters_g}}
    \begin{aligned}
        g(n, e) = \frac{n^{4}}{32} & \Big\{ \Big[ J_{n-2}(n e)-2 e J_{n-1}(n e)+\frac{2}{n} J_{n}(n e) \\
        &\left. +2 e J_{n+1}(n e)-J_{n+2}(n e)\Big]^{2}\right.\\
        &\left.\kern-\nulldelimiterspace\right.
            \!\!\begin{aligned}
            +\big(1-e^{2}\big)\Big[&J_{n-2}(n e)-2 J_{n}(n e) \\
            &+J_{n+2}(n e)\Big]^{2}
            \end{aligned}\\
        &+\frac{4}{3 n^{2}}\Big[J_{n}(n e)\Big]^{2}\Big\},
    \end{aligned}
    \label{eq:g(n,e)}
\end{equation}
where $J_{n}(v)$ is the Bessel function of the first kind. Thus, the \textit{sum} of $g(n, e)$ over all harmonics gives the factor by which the gravitational-wave emission is stronger for a binary of eccentricity $e$ over an equivalent circular binary. This enhancement factor is \citep[][Eq.\,17]{Peters+1963}
\begin{equation}{\lwFuncURL{utils}{peters_f}}
    F(e) \equiv \sum_{n = 1}^{\infty} g(n, e) = \frac{1 + \frac{73}{24} e^2 + \frac{37}{96} e^4}{(1 - e^2)^{7/2}}.
    \label{eq:eccentricity_enhancement_factor}
\end{equation}
A useful rule of thumb is that $F(0.5) \approx 5.0$, or in words, a binary with eccentricity $0.5$ loses energy to gravitational waves at 5 times the rate of an equivalent circular binary.

@@

The characteristic strain from a stellar mass binary in the $n^{\rm th}$ harmonic is given by (e.g. \citealp[][Eq.\,56]{Barack+2004}; \citealp[][Eq.\,5.1]{ Flanagan+1998})
\begin{equation}{}
    h_{c,n}^2 = \frac{1}{(\pi D_L)^2} \qty( \frac{2 G}{c^3} \frac{\dot{E_n}}{\dot{f_n}} ),
    \label{eq:char_strain_dedf}
\end{equation}
where $D_L$ is the luminosity distance to the source (note that for Milky Way sources this is simply the distance to the source), $\dot{E}_n$ is the power radiated in the $n^{\rm th}$ harmonic and $\dot{f}_n$ is the rate of change of the $n^{\rm th}$ harmonic frequency.

@@

\subsubsection{TianQin}
For the TianQin sensitivity curve we use the power spectral density given in \citet[][Eq.~13]{Huang+2020}
\begin{equation}{}
    \begin{split}
    S_{N}(f) &= \frac{1}{L^{2}}\left[\frac{4 S_{a}}{(2 \pi f)^{4}}\left(1+\frac{10^{-4} \unit{Hz}}{f}\right)+S_{x}\right] \\
    & \times\left[1+0.6\left(\frac{f}{f_{*}}\right)^{2}\right],
    \end{split}
\label{eq:tianqin}
\end{equation}
where $L = \sqrt{3} \times 10^5 \unit{km}$ is the arm length, $S_a = 1 \times 10^{-30} \unit{m^2}{s^{-4}}{Hz^{-1}}$ is the acceleration noise, $S_x = 1 \times 10^{-24} \unit{m^2}{Hz^{-1}}$ is the displacement measurement noise and $f_* = c / 2 \pi L$ is the transfer frequency.

@@

\begin{acknowledgements}
    % People
    We are grateful to Stat Babak, Floor Broekgaarden, Tom Callister, Will Farr, Stephen Justham, Valeria Korol, Mike Lau, Tyson Littenberg, Ilya Mandel, Alberto Sesana, Lieke van Son, the CCA GW group, the BinCosmos group and the COMPAS group for stimulating discussions that influenced and motivated us to complete this project. We thank the BinCosmos group for testing an early version of the package and providing useful feedback. In particular, we thank Lieke van Son for her innovation in inventing the name \lw{}! TW also thanks Floor Broekgaarden for first suggesting that he investigate LISA and the derivation of the SNR calculation.

>>>>

% LEGWORK stuff
\newcommand{\lw}{LEGWORK}
\newcommand{\lwColour}{SeaGreen}
\newcommand{\lwModURL}[1]{https://legwork.readthedocs.io/en/latest/modules.html\#module-legwork.#1}
\newcommand{\lwModLink}[1]{\href{https://legwork.readthedocs.io/en/latest/modules.html\#module-legwork.#1}{\color{\lwColour}{\texttt{#1}}}}
\newcommand{\lwFuncURL}[2]{https://legwork.readthedocs.io/en/latest/api/legwork.#1.#2.html}

@@

It is often convenient to convert between orbital frequency, $f_{\rm orb}$, and the semi-major axis, $a$, of a binary and this can be accomplished with Kepler's third law.
\begin{equation}{\lwFuncURL{utils}{get_a_from_f_orb}}
    a = \qty(\frac{G(m_1 + m_2)}{(2 \pi f_{\rm orb})^{2}})^{1/3},
    \label{eq:kepler3rd}
\end{equation}
where $G$ is the gravitational constant. Inversely we could write
\begin{equation}{\lwFuncURL{utils}{get_a_from_f_orb}}
    f_{\rm orb} = \frac{1}{2 \pi} \sqrt{\frac{G(m_1 + m_2)}{a^3}}.
    \label{eq:kepler3rd_reverse}
\end{equation}
For circular binaries, gravitational-wave emission occurs at twice the orbital frequency ($f_{\rm GW} = 2 f_{\rm orb}$). However, for eccentric binaries, we need to consider all frequency harmonics of gravitational-wave emission. These are defined such that the $n^{\rm th}$ harmonic frequency is
\begin{equation}{}
    f_{n} \equiv n \cdot f_{\rm orb}.
    \label{eq:nth_freq}
\end{equation}
It will be important to know the relative gravitational-wave power radiated into the $n^{\rm th}$ harmonic for a binary with eccentricity $e$ for the following strain and SNR calculations. This is given by \citet[][Eq.\,20]{Peters+1963}
\begin{equation}{\lwFuncURL{utils}{peters_g}}
    \begin{aligned}
        g(n, e) = \frac{n^{4}}{32} & \Big\{ \Big[ J_{n-2}(n e)-2 e J_{n-1}(n e)+\frac{2}{n} J_{n}(n e) \\
        &\left. +2 e J_{n+1}(n e)-J_{n+2}(n e)\Big]^{2}\right.\\
        &\left.\kern-\nulldelimiterspace\right.
            \!\!\begin{aligned}
            +\big(1-e^{2}\big)\Big[&J_{n-2}(n e)-2 J_{n}(n e) \\
            &+J_{n+2}(n e)\Big]^{2}
            \end{aligned}\\
        &+\frac{4}{3 n^{2}}\Big[J_{n}(n e)\Big]^{2}\Big\},
    \end{aligned}
    \label{eq:g(n,e)}
\end{equation}
where $J_{n}(v)$ is the Bessel function of the first kind. Thus, the \textit{sum} of $g(n, e)$ over all harmonics gives the factor by which the gravitational-wave emission is stronger for a binary of eccentricity $e$ over an equivalent circular binary. This enhancement factor is \citep[][Eq.\,17]{Peters+1963}
\begin{equation}{\lwFuncURL{utils}{peters_f}}
    F(e) \equiv \sum_{n = 1}^{\infty} g(n, e) = \frac{1 + \frac{73}{24} e^2 + \frac{37}{96} e^4}{(1 - e^2)^{7/2}}.
    \label{eq:eccentricity_enhancement_factor}
\end{equation}
A useful rule of thumb is that $F(0.5) \approx 5.0$, or in words, a binary with eccentricity $0.5$ loses energy to gravitational waves at 5 times the rate of an equivalent circular binary.

@@

The characteristic strain from a stellar mass binary in the $n^{\rm th}$ harmonic is given by (e.g. \citealp[][Eq.\,56]{Barack+2004}; \citealp[][Eq.\,5.1]{ Flanagan+1998})
\begin{equation}{}
    h_{c,n}^2 = \frac{1}{(\pi D_L)^2} \qty( \frac{2 G}{c^3} \frac{\dot{E_n}}{\dot{f_n}} ),
    \label{eq:char_strain_dedf}
\end{equation}
where $D_L$ is the luminosity distance to the source (note that for Milky Way sources, or any sources with redshift $\sim 0$, this is simply the distance to the source), $\dot{E}_n$ is the power radiated in the $n^{\rm th}$ harmonic and $\dot{f}_n$ is the rate of change of the $n^{\rm th}$ harmonic frequency.

@@

\subsubsection{TianQin}
For the TianQin sensitivity curve we use the power spectral density given in \citet[][Eq.~13]{Huang+2020}
\begin{equation}{\lwFuncURL{psd}{tianqin_psd}}
    \begin{split}
    S_{N}(f) &= \frac{1}{L^{2}}\left[\frac{4 S_{a}}{(2 \pi f)^{4}}\left(1+\frac{10^{-4} \unit{Hz}}{f}\right)+S_{x}\right] \\
    & \times\left[1+0.6\left(\frac{f}{f_{*}}\right)^{2}\right],
    \end{split}
\label{eq:tianqin}
\end{equation}
where $L = \sqrt{3} \times 10^5 \unit{km}$ is the arm length, $S_a = 1 \times 10^{-30} \unit{m^2}{s^{-4}}{Hz^{-1}}$ is the acceleration noise, $S_x = 1 \times 10^{-24} \unit{m^2}{Hz^{-1}}$ is the displacement measurement noise and $f_* = c / 2 \pi L$ is the transfer frequency.

@@

\begin{acknowledgements}
    % People
    We are grateful to Stas Babak, Floor Broekgaarden, Tom Callister, Will Farr, Stephen Justham, Valeria Korol, Mike Lau, Tyson Littenberg, Ilya Mandel, Alberto Sesana, Lieke van Son, the CCA GW group, the BinCosmos group and the COMPAS group for stimulating discussions that influenced and motivated us to complete this project. We thank the BinCosmos group for testing an early version of the package and providing useful feedback. In particular, we thank Lieke van Son for her innovation in inventing the name \lw{}! TW also thanks Floor Broekgaarden for first suggesting that he investigate LISA and the derivation of the SNR calculation.
